---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FactClust

<!-- badges: start -->
<!-- badges: end -->

The goal of FactClust is to simplify the clustering algorithm for factor variables with many levels when the data needs to be rolled up.

## Installation

You can install the development version of FactClust from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("easoneli176/FactClust")
```

## Example

When we have a factor variable with many levels, and we need to rollup the data by some grouping factor and calculate clusters, we begin by one-hot encoding the factor variable using the onehot function:

```{r example}
library(FacClust)

mock_data <- data.frame(fac = c(rep("A",10),rep("B",10),rep("C",10),rep("D",10),rep("E",10),rep("F",10)),grp=c(rep("G1",15),rep("G2",15),rep("G3",15),rep("G4",15)))

onehot(mock_data$fac)
```

Next, we calculate proportions of each factor level within each group. We can optionally use a weighting variable here to make the proportions weighted.

```{r example2}
df<-PropRollup(mock_data$fac,mock_data$grp)
df
```

Next, we can calculate clusters on these proportions using kmeans from the base R package. The number of clusters, k, should be determined outside this package.

```{r example3}
set.seed(278613)
km<-kmeans(x=df[,-1],centers=2)
km
```

Finally, as long as we have named our kmeans object km, we can assign clusters to out of sample data using this function:

```{r example4}
apply(df[,-1], 1, oos_clust)
```
